\vspace{-1cm} \normalsize Supplementary Information\\\vspace{0.2cm} \LARGE Basic aspects of species' distributions

### 2021-06-22 manuscript/manuscript.tex
simulations for baseline

\subsubsection*{Distribution data}
We used data describing the distribution of 798 species across 912 sites covering most of the mountain region of the Western Alps in the Canton de Vaud (Switzerland; \citealt{scherrerEcologicalIndicatorValues2019}). Each of these sites is a $8\times 8\,\text{m}$ plot placed somewhere along an elevation range from $375\,\text{m}$ to $3210\,\text{m}$. In all sites, presence/absence data as well as Braun-Blanquet abundance-dominance classes were recorded for all species. Additionally, we used meteorological data provided by \citet{scherrerEcologicalIndicatorValues2019}, containing multiple variables characterizing the climate in each site at high spatial resolution ($25\,\text{m}$). This dataset was compiled based on 30 years (1961–1990) of records from national weather stations. Since most of the data is highly correlated, we calculated the main axes of variation of the following scaled variables: daily minimum, maximum and average temperature; sum of growing degree-days above $5^{\circ}\text{C}$; mean temperature of wettest quarter; annual precipitation, precipitation seasonality, and precipitation of driest quarter (see Supplementary Methods; Supplementary Fig. 1).  %Missing TabsY

@@

Species' environmental preferences in this database can be used to inform distribution models---e.g.~as an informative prior in a Bayesian framework. These are characterized following the ecological indicator values developed by \citet{landoltFloraIndicativaOkologische2010}, providing both an estimate of the average conditions in which a species can be found as well as a broad description of their range of variation. These values are provided for a range of 10 environmental variables, including temperature, continentality, light conditions, as well as moisture, acidity and nutrient content of the soil (see a full list and description of the ecological indicators in the Supplementary Table 1; \citealt{landoltFloraIndicativaOkologische2010}). In addition to species' environmental preferences, the floristic data also contains information on species introduction status (e.g.~identifying those species that are recent and historical range expanders) and change tendency (e.g.~indicating species that have shown decline or increase in their populations over the recent decades). We describe this information in more detail in the Supplementary Table 1.

@@

To test the performance of the model as well as our choice of prior distributions, we modelled simulated data and compared the sampled posterior distributions to the data-generating parameters (see Supplementary Methods; Supplementary Fig. 2). Notice that using the link function in Eq.~(\ref{eq:baseline-response}) could cause problems when sampling the model, and some adjustments need to be made when specifying the model (see Supplementary Methods and the Code Availability section). To perform the data analysis and generate the figures, we used some of the functions available with the R package `rethinking' \citep{mcelreathStatisticalRethinkingBayesian2020}.

@@

\subsubsection*{Fat-tailed response curve}
Fat-tailed distributions represent distributions with relatively high representation of extreme events. While many different distributions exhibit this property, we decided to accommodate this feature into our baseline model by considering a response curve that follows a generalized error distribution. Such a distribution is useful because the Gaussian shape is a special case of it, and it contains a parameter that regulates the level of kurtosis---ranging from longer to shorter tails than the Gaussian case (Fig~\ref{fig:response}b). In particular, we can adapt Eq.~(\ref{eq:baseline-response}) to present this non-linear form as follows:
\begin{equation}
\begin{split}
\text{log}\left(p_{ij}\right) = -\alpha_{i} - \gamma^{\prime}_{i}\, |x_{j}-\beta_{i}|^{\nu_{i}} ,
\end{split}
\label{eq:generalized-response}
\end{equation}
where $\gamma^{\prime}_{i} = g(\gamma_{i}, \nu_{i})$, and $\nu_{i}$ is a parameter that describes the kurtosis of the distribution, which we define as $\nu_{i}\in\left(1, \infty\right)$. Following this, we choose an adaptive prior for this set of new parameter such that $\log\left(\nu_{i}-1\right)\sim \text{Normal}\left(\hat{\nu}, \sigma_{\nu}\right)$, where $\hat{\nu}\sim\text{Normal}\left(0, 1\right)$ and $\sigma_{\nu}\sim\text{Exponential}\left(2\right)$. Given the relationship between $\gamma^{\prime}_{i}$ and $\gamma_{i}$,  we can re-parametrize the model and follow Eq.~(\ref{eq:baseline-priors}) to define the prior distributions (see Supplementary Table 2; Code Availability section; \citealt{nadarajahGeneralizedNormalDistribution2005}). Notice that the Gaussian distribution will naturally emerge when $\nu_i=2$.

@@

\subsubsection*{Skewed response curve}
Skewed responses present steeper declines towards either side of the distribution. One way to accommodate this feature in our models is by considering a skewed normal distribution (Supplementary Figure XX). As for the case described above, the Gaussian is a special case of this distribution, and it contains a parameter that controls for the level and direction of `skewness' (Fig~\ref{fig:response}c). Importantly, this distribution presents normal-like tails; therefore, the added skewness does not make additional assumptions regarding how species are distributed along the gradient. To test for the existence of this feature, we modified Eq.~(\ref{eq:baseline-response}) as
\begin{equation}	
\begin{split}
\text{log}\left(p_{ij}\right) = -\alpha_{i} - \gamma^{\prime}_{i}\left(\frac{x_{j}-\beta^{\prime}_{i}}{1+\lambda_{i}\,\text{sgn}\left(x_{j}-\beta^{\prime}_{i}\right)}\right)^{2} ,
\end{split}
\label{eq:skewed-response}
\end{equation}
where $\gamma^{\prime}_{i} = q_{1}\left(\gamma_{i}, \nu_{i}, \lambda_{i}\right)$, $\beta^{\prime}_{i} = q_{2}\left(\gamma_{i}, \beta_{i}, \nu_{i}, \lambda_{i}\right)$, and $\lambda_{i}$ is a parameter that describes the skewness of the distribution such that $\lambda_{i}\in\left(-1, 1\right)$. The function $\text{sgn}\left(x\right)$ characterizes the sign function.  We chose $\lambda_i$ to have an adaptive prior such that $\text{logit}\left(\frac{\lambda_{i}+1}{2}\right)\sim \text{Normal}\left(\hat{\lambda}, \sigma_{\lambda}\right)$, where $\hat{\lambda}\sim\text{Normal}\left(0, 1\right)$ and $\sigma_{\lambda}\sim\text{Exponential}\left(1\right)$. Notice that this model can be re-parametrized following $q_1$ and $q_2$, allowing us to set the rest of the prior distributions as described for the baseline model (see Supplementary Table 2; Code Availability section). In this case, the Gaussian distribution is a special case of Eq.~(\ref{eq:skewed-response}) when $\lambda_{i}=0$ \citep{ashourApproximateSkewNormal2010}.

>>>>

\subsubsection*{Distribution data}
We used data describing the distribution of 798 species across 912 sites covering most of the mountain region of the Western Alps in the Canton de Vaud (Switzerland; \citealt{scherrerEcologicalIndicatorValues2019}). Each of these sites is a $8\times 8\,\text{m}$ plot placed somewhere along an elevation range from $375\,\text{m}$ to $3210\,\text{m}$. In all sites, presence/absence data as well as Braun-Blanquet abundance-dominance classes were recorded for all species. Additionally, we used meteorological data provided by \citet{scherrerEcologicalIndicatorValues2019}, containing multiple variables characterizing the climate in each site at high spatial resolution ($25\,\text{m}$). This dataset was compiled based on 30 years (1961–1990) of records from national weather stations. Since most of the data is highly correlated, we calculated the main axes of variation of the following scaled variables: daily minimum, maximum and average temperature; sum of growing degree-days above $5^{\circ}\text{C}$; mean temperature of wettest quarter; annual precipitation, precipitation seasonality, and precipitation of driest quarter (Supplementary Fig.~1).  %Missing TabsY

@@

Species' environmental preferences in this database can be used to inform distribution models---e.g.~as an informative prior in a Bayesian framework. These are characterized following the ecological indicator values developed by \citet{landoltFloraIndicativaOkologische2010}, providing both an estimate of the average conditions in which a species can be found as well as a broad description of their range of variation. These values are provided for a range of 8 environmental variables, including temperature, continentality, light conditions, as well as moisture, acidity and nutrient content of the soil (see a full list and description of the ecological indicators in the Supplementary Table 1; \citealt{landoltFloraIndicativaOkologische2010}). In addition to species' environmental preferences, the floristic data also contains information on species introduction status (e.g.~identifying those species that are recent and historical range expanders) and change tendency (e.g.~indicating species that have shown decline or increase in their populations over the recent decades). We describe this information in more detail in the Supplementary Table 1.

@@

To test the performance of the model as well as our choice of prior distributions, we modelled simulated data and compared the sampled posterior distributions to the data-generating parameters (see Code Availability section). Notice that using the link function in Eq.~(\ref{eq:baseline-response}) could cause problems when sampling the model, and some adjustments need to be made when specifying the model (see Code Availability section). To perform the data analysis and generate the figures, we used some of the functions available with the R package `rethinking' \citep{mcelreathStatisticalRethinkingBayesian2020}.

@@

\subsubsection*{Fat-tailed response curve}
Fat-tailed distributions represent distributions with relatively high representation of extreme events. While many different distributions exhibit this property, we decided to accommodate this feature into our baseline model by considering a response curve that follows a generalized error distribution. Such a distribution is useful because the Gaussian shape is a special case of it, and it contains a parameter that regulates the level of kurtosis---ranging from longer to shorter tails than the Gaussian case (Fig~\ref{fig:response}b). In particular, we can adapt Eq.~(\ref{eq:baseline-response}) to present this non-linear form as follows:
\begin{equation}
\begin{split}
\text{log}\left(p_{ij}\right) = -\alpha_{i} - \gamma^{\prime}_{i}\, |x_{j}-\beta_{i}|^{\nu_{i}} ,
\end{split}
\label{eq:generalized-response}
\end{equation}
where $\gamma^{\prime}_{i} = g(\gamma_{i}, \nu_{i})$, and $\nu_{i}$ is a parameter that describes the kurtosis of the distribution, which we define as $\nu_{i}\in\left(1, \infty\right)$. Following this, we choose an adaptive prior for this set of new parameter such that $\log\left(\nu_{i}-1\right)\sim \text{Normal}\left(\hat{\nu}, \sigma_{\nu}\right)$, where $\hat{\nu}\sim\text{Normal}\left(0, 1\right)$ and $\sigma_{\nu}\sim\text{Exponential}\left(2\right)$. Given the relationship between $\gamma^{\prime}_{i}$ and $\gamma_{i}$,  we can re-parametrize the model and follow Eq.~(\ref{eq:baseline-priors}) to define the prior distributions (see Supplementary Table 2; \citealt{nadarajahGeneralizedNormalDistribution2005}). Notice that the Gaussian distribution will naturally emerge when $\nu_i=2$.

@@

\subsubsection*{Skewed response curve}
Skewed responses present steeper declines towards either side of the distribution. One way to accommodate this feature in our models is by considering a skewed normal distribution. As for the case described above, the Gaussian is a special case of this distribution, and it contains a parameter that controls for the level and direction of `skewness' (Fig~\ref{fig:response}c). Importantly, this distribution presents normal-like tails; therefore, the added skewness does not make additional assumptions regarding how species are distributed along the gradient. To test for the existence of this feature, we modified Eq.~(\ref{eq:baseline-response}) as
\begin{equation}	
\begin{split}
\text{log}\left(p_{ij}\right) = -\alpha_{i} - \gamma^{\prime}_{i}\left(\frac{x_{j}-\beta^{\prime}_{i}}{1+\lambda_{i}\,\text{sgn}\left(x_{j}-\beta^{\prime}_{i}\right)}\right)^{2} ,
\end{split}
\label{eq:skewed-response}
\end{equation}
where $\gamma^{\prime}_{i} = q_{1}\left(\gamma_{i}, \lambda_{i}\right)$, $\beta^{\prime}_{i} = q_{2}\left(\gamma_{i}, \beta_{i}, \lambda_{i}\right)$, and $\lambda_{i}$ is a parameter that describes the skewness of the distribution such that $\lambda_{i}\in\left(-1, 1\right)$. The function $\text{sgn}\left(x\right)$ characterizes the sign function.  We chose $\lambda_i$ to have an adaptive prior such that $\text{logit}\left(\frac{\lambda_{i}+1}{2}\right)\sim \text{Normal}\left(\hat{\lambda}, \sigma_{\lambda}\right)$, where $\hat{\lambda}\sim\text{Normal}\left(0, 1\right)$ and $\sigma_{\lambda}\sim\text{Exponential}\left(1\right)$. Notice that this model can be re-parametrized following $q_1$ and $q_2$, allowing us to set the rest of the prior distributions as described for the baseline model (see Supplementary Table 2; Code Availability section). In this case, the Gaussian distribution is a special case of Eq.~(\ref{eq:skewed-response}) when $\lambda_{i}=0$ \citep{ashourApproximateSkewNormal2010}.

### 2021-06-22 manuscript/supplementary_information.tex
simulations for baseline

% Maths tools
\usepackage{amsmath}
\usepackage{amssymb}

@@

@@

@@

@@

\author{\textit{Bramon Mora and Alexander}}

@@

@@

@@

\bibliography{references_ab}

>>>>

% Maths tools
\usepackage[tbtags]{amsmath}
\usepackage{amssymb}

@@

% Landscape page
\usepackage{pdflscape}
\usepackage{afterpage}

@@

\usepackage{makecell}

@@

\renewcommand{\arraystretch}{1.3}

@@

\author{\textit{Bramon Mora et.\,al.}}

@@

% Full description of Binomial model with multiple environmental covariates
% Full description of Categorical model with multiple environmental covariates
% Full description of Categorical model with heavy-tails---generalized error distribution
% Full description of Categorical model with heavy-tails---Student's t-distribution
% Full description of Categorical model with skewed---skewed normal distribution
% Using a MMSBM to calculate species similarity
% Edge problem in stan and need to define minimum probability.

\section*{Supplementary Methods}
\subsubsection*{Full description of the baseline model}
While the baseline model described in the main text included only one predictor, the same model can be extended to include multiple environmental variables. For example, imagine that we had $k$ predictors describing the environmental conditions of any given site $j$. The baseline model in Eq.~(1) of the main text could then be rewritten as:
\begin{equation}
\begin{split}
y_{ij} & \sim \text{Binomial}\left(1, p_{ij}\right) \\
\text{log}\left(p_{ij}\right) & = -\alpha_{i} - \sum_k \gamma_{ik} \left(x_{jk}-\beta_{ik}\right)^2,
\end{split}
\label{eq:baseline-binary}
\end{equation}
where $y_ij$ describes the presence and absence of species, and the prior distributions for the different parameters are analogous to the ones described in the main text.

Similarly, if the response variable $y_ij$ represented ordinal data (e.g.~Braun-Blanquet abundance-dominance classes), characterized by $K$ different classes. Then, one can adapt Supplementary Eq.~(\ref{eq:baseline-binary}) as:
\begin{equation}
\begin{split}
y_{ij} & \sim \text{Categorical}\left(\mathbf{q}\right) \\
q_{1} & = 1- p_{ij}\phi_1 \\
q_{k} & =  p_{ij}\phi_{k-1} -  p_{ij}\phi_{k} \\
q_{K} & =  p_{ij}\phi_{K-1} \\
\text{log}\left(p_{ij}\right) & = -\alpha_{i} - \sum_k \gamma_{ik} \left(x_{jk}-\beta_{ik}\right)^2 \\
\phi_1 & = \hat{\phi_1} \\
\phi_k & = \hat{\phi_k} \,\, \hat{\phi_{k-1}} \\
\hat{\phi} & = \text{Beta}(1,1),
\end{split}
\label{eq:baseline-categorical}
\end{equation}
where the prior distributions for the rest of parameters are analogous to the ones described in the main text.

\subsubsection*{Distance matrices from incomplete categorical and ordinal data}
The prior information that we have regarding species' distributions is represented by the set of ordinal and categorical traits found in the floristic database. More specifically, both the ecological indicator values and range of variation are ordinal traits, containing potential missing entries (Supplementary Table \ref{stab:EI}). These data could be directly used as covariates in any given distribution model; however, we want this information to be accounted for as a prior for the relevant parameters of our Bayesian model. One way to do so is by compiling the traits in the floristic database into variance-covariance matrices characterizing the \textit{a priori} similarity between species. To accomplish this, we need to turn the set of ecological indicator values and range of variation into the distance matrices $D$ presented in the text.
%, whereas plants' physiological data are characterized by categorical data containing multiple missing entries.

More generally, we want to understand the way $N$ species are characterized by $M$ categorical traits. One way to frame this problem is by using a network representation. Following the ideas presented by \citet{godoy-loriteAccurateScalableSocial2016}, we assume that species can be connected to each of these traits by an interaction $\left(i, j\right)$ that can be of any type $r\in R$. Notice that this provides us with multiple ways to account for the information---and lack thereof---contained in the different ordinal traits $M$. That is, the $R$ types of interactions can represent the lack of information for a particular link $\left(i, j\right)$, the absence or presence of such interaction, and any type of association between $i$ and $j$. Then, given a set of interactions $R^{*}$ between $N$ and $M$, we can use a Mixed Membership Stochastic Block Model (MMSBM) to understand the differences between species. In particular, these models consider that the $N$ plants and $M$ traits can be classified into $K$ and $L$ groups types, respectively. Then, they estimate `group-membership vectors', describing the extend to which the different plants and traits belong to the different group types.

Here, we use the MMSBM presented by \citet{godoy-loriteAccurateScalableSocial2016}, and classify plants into 5 types of species, and traits into 5 trait groups. Then, we estimate the group-membership vectors $\vec{\theta}_{\text{plants}}$, and calculate the pairwise distances $D_{ij}$ between plants as the euclidean distance between their membership vectors.  We run these models independently for species' ecological indicator values and their range of variation (Supplementary Table \ref{stab:EI}) to generate the two distance matrices for parameters $\beta_i$ and $\gamma_i$. Notice that the MMSBM uses an expectation-maximization algorithm to estimate the different membership vectors. Therefore, in order to have good estimates of the different pairwise distances $D_{ij}$ for $\beta_i$ and $\gamma_i$, one needs to run the MMSBM several times (1000 times) and average the resulting pairwise distances \citep{godoy-loriteAccurateScalableSocial2016, tarres-deulofeuTensorialBipartiteBlock2019}.

This type of classification of ordinal variables that describe different species can also be done with other algorithms, including categorical PCA and matrix factorization methods. However, the MMSBM provides a convenient tool to classify very different types of traits, including categorical variables or diverse data containing multiple missing entries. Moreover, MMSBM have been shown to better capture the patterns of this sort of data \citet{godoy-loriteAccurateScalableSocial2016}.

%Given a set of interactions $R^{*}$ between $N$ and $M$, we use a Mixed Membership Stochastic Block Model (MMSBM) to characterize these. In particular, we consider that plants and traits can be classified into $K$ and $L$ groups, respectively. For every species $i$, we assume that there is a probability $\theta_{i\alpha}$ for it to belong to any of the $K$ species groups. Likewise, we also assume that any trait $j$ has a probability $\phi_{j\beta}$ of belonging to any of the $L$ trait groups. Finally, we define $p_{\alpha\beta}\left(r\right)$ as the probability of a species from group $\alpha$ interacting with a trait from group $\beta$ by an association type $r$. Putting these together, the probability of an interaction $\left(i, j\right)$ of type $r$ can be calculated as:
%\begin{equation}
%Pr[r_{ij}=r] = \sum_{\alpha \beta} \theta_{i\alpha} \phi_{j\beta} p_{\alpha\beta}\left(r\right)
%\end{equation}
%Following this definition, we want to find the group memberships that maximize the likelihood $P\left(R^{*}|\theta, \phi, p\right)$. Doing so is difficult optimization problem; however, it has been shown that one can estimate the different $\theta_{i\alpha}$, $\phi_{j\beta}$, and $p_{\alpha\beta}\left(r\right)$ parameters by maximizing the likelihood using an expectation-maximization algorithm \citep{godoy-loriteAccurateScalableSocial2016, tarres-deulofeuTensorialBipartiteBlock2019}. In simple terms, one can iteratively find multiple local minima for the likelihood, and average over the estimated the parameter values \citep{godoy-loriteAccurateScalableSocial2016}\footnote[2]{
%While this averaging is trivial for the estimated probabilities $Pr[r_{ij}=r]$, it is non-trivial if one wants to find averages for the group memberships. The reason for this is related to the stochastic nature of the expectation-maximization algorithm. This algorithm initially assigns random group memberships to both species and traits. While this random labelling is irrelevant when studying the probabilities $Pr[r_{ij}=r]$, it is instead crucial for averaging $\theta_{i\alpha}$, $\phi_{j\beta}$, and $p_{\alpha\beta}\left(r\right)$. Therefore, before averaging the group membership estimates, one needs to find the bijective relationship for the labellings of different iterations of the optimization algorithm. In a nutshell, for every iteration, I do this by using a simulated annealing algorithm on the estimated $p_{\alpha\beta}\left(r\right)$, matching the corresponding labelling to a reference iteration.}.

%The average estimates for the group memberships provide us with a different scale to classify species based on the traits these have. In short, for any species $i$, we can estimate a $K$-dimensional vector $\vec{\theta}_{i}$ that describes the extend to which $i$ belong to each group membership---i.e. the extend to which a species is of one type or another. This classification is useful because it can be used to compare species, defining a way to measure the distance between species based on an arbitrary---and potentially incomplete---set of categorical or ordinal traits $M$. The simplest case is to define the distance as $D_{ij} = |\vec{\theta}_{i}-\vec{\theta}_{j}|$. Alternatively, one could also define $K$ distance matrices based on the different group memberships $D^{\alpha}_{ij} = |\theta_{i\alpha}-\theta_{j\alpha}|$.

%\subsubsection*{Sampling the posterior: simulated data and additional information}
%A good way to understand the model behaviour and the choice of prior distributions is by using simulated data. Here, we only present the simulations for the baseline model---the simplest model---but similar simulations were done for the others (see Code Availability section of the main text). 
%
%To simulate the data, we first generated two distance matrices relating the different parameters

\subsubsection*{Alternative variance-covariance structures}
The model structure defined above allows us to test how different sources of information characterize each of the different parameters. Specifically, we can do this by modifying Eq.~(\ref{eq:covariance-baseline}). For example, imagine that we have multiple matrices $D^k$ characterizing species' differences along different axis of variation---e.g.~two matrices characterizing physiological and environmental traits. One can modify Eq.~(\ref{eq:covariance-baseline}) for a particular parameter---e.g.~$\beta_{i}$---such that
\begin{equation} 
\Sigma_{ij} = \eta\,\text{exp}\left(-\sum_k\rho_{k} {D^{k}_{ij}}^2\right) + \delta_{ij} \sigma ,
\label{eq:covariance-complex}
\end{equation}
where now $\rho_{k}$ are separate relevance hyperparameters for each distance matrix in the total variance of $\beta_i$.

\section*{Supplementary Notes}
\section*{Supplementary Figures}

\begin{figure}[ht]
  \centering
    \vspace{0.5cm}
    \includegraphics[width=1\textwidth]{figures/variances}
    	  \vspace{0.3cm}
	   \caption{PCA and variance explained.}
      \label{sfig:pca}
\end{figure}

@@

\begin{figure}[ht]
  \centering
    \vspace{0.5cm}
    \includegraphics[width=1\textwidth]{figures/pca-elevation-temperature}
    	  \vspace{0.3cm}
	   \caption{The same as in Fig.~1 but with elevation and temperature.}
      \label{sfig:temperature-elevation}
\end{figure}

@@

\section*{Supplementary Tables}

\begin{table}[h]
\begin{center}
\begin{tabular}{ p{8em} p{6em} l p{16em}} 
\textbf{Information} & \textbf{Variable type} & \textbf{Variable} & \textbf{Description} \\
\hline
\multirow{3}{7em}{ecological indicator values} & \multirow{3}{6em}{climate indicators} & temperature & \textit{10 classes ranging from alpine climates to very warm collines} \\
& & continentality & \textit{5 classes ranging from oceanic to continental climates}\\ 
& & light & \textit{5 classes ranging from deep shade to full light}\\ 
\cline{2-4}
& \multirow{5}{6em}{soil indicators} & moisture & \textit{10 classes ranging from very dry to flooded soil} \\
& & reaction & \textit{5 classes ranging from extremely acid to alkaline soil} \\
& & nutrients & \textit{5 classes ranging from very infertile to over-rich soils} \\ 
& & humus & \textit{5 classes ranging from little to high humus content} \\
& & aeration & \textit{5 classes ranging from bad to good aeration} \\ 
\hline
\multirow{3}{7em}{range of variation} & \multirow{3}{6em}{climate indicators} & temperature & \multirow{3}{14em}{\textit{3 classes indicating small, large and unlimited variation of the different climatic indicators}} \\  
& & continentality & \\ 
& & light & \\ 
\cline{2-4}
& \multirow{3}{6em}{soil indicators} & moisture & \multirow{5}{14em}{\textit{3 classes indicating small, large and unlimited variation of the different soil indicators}}  \\ 
& & reaction & \\ 
& & nutrients & \\ 
& & humus & \\ 
& & aeration & \\ 
\hline
\multirow{3}{7em}{species types} & \multirow{2}{6em}{time of immigration} & invasion status & \textit{classified as indigenous, historical range expanders and recent range expanders} \\  
& \makecell[l]{ \\change\\tendency} & abundance & \textit{classified as stable, decreasing, decreasing at low elevations, and increasing in abundance}\\ 
\hline  
\end{tabular}
\end{center}
\caption{Indicator values \citep{landoltFloraIndicativaOkologische2010}}
\label{stab:EI}   
\end{table}

\begin{table}[h]
\begin{center}
\begin{tabular}{  p{6em} | p{1em} l l } 
\textbf{response curve} & &  \(\beta^{\prime}_i\) & \(\gamma^{\prime}_i\) \\
\\[-1em]
\hline
\\[-1em]
baseline & & \(\beta_i\) & \(\gamma_i\) \\
\\[-1em]

\\[-1em]
fat-tailed & & \(\beta_i\) & \( g = \sqrt{\frac{\gamma_i \Gamma\left(\frac{3}{\nu_i}\right)}{\Gamma\left(\frac{1}{\nu_i}\right)}}\) \\
\\[-1em]

\\[-1em]
skewed &  & \( q_2 = \beta_i - \frac{2 \lambda_i}{\sqrt{\pi} \gamma^{\prime}_i}  \)  & \( q_1 = \gamma_i  \sqrt{\frac{\pi\left(1+3 \lambda^{2}_{i}\right)-8\lambda^2_i}{2\pi}}  \)  \\
\\[-1em]

\\[-1em]
\multirow{-1.25}{6em}{skewed and fat-tailed} & & \( f_2 = \beta_i - \frac{2^{\frac{2}{\nu_i}}\lambda_i \Gamma\left(\frac{1}{2}+\frac{1}{\nu_i}\right)}{\sqrt{\pi}\gamma^{\prime}_i}   \)  & \( f_1 = \gamma_i \sqrt{\frac{\pi \left(1+3\lambda^2_i\right) \Gamma\left(\frac{3}{\nu_i}\right)-16^{\frac{1}{\nu_i}} \lambda^{2}_{i} \Gamma\left(\frac{1}{2}+\frac{1}{\nu_i}\right)^2 \Gamma\left(\frac{1}{\nu_i}\right)}{\pi \Gamma\left(\frac{1}{\nu_i}\right)}} \) \\
\\[-1em]

\end{tabular}
\end{center}
\caption{cite vignette}
\label{stab:formulas}   
\end{table}

\bibliography{references2}
